---
title: "SwiftKey NLP Milestone Report"
author: "Tianxiang(Ivan) Liu"
date: "13 November 2014"
output: html_document
---

```{r echo=FALSE, message=FALSE}
setwd('/Users/ivan/Work_directory/SwiftKey')
require(tm); require(SnowballC); require(data.table)
require(ggplot2); require(RWeka); require(qdap);
require(scales); require(gridExtra); require(wordcloud);require(reshape2)
rm(list=ls(all=TRUE));gc(reset=TRUE);par(mfrow=c(1,1))
source('SwiftKey-Natural-language/Task_1.5_Tokenization_func.R')
```


### Introduction
This milestone report introduced a preliminary research about SwiftKey's NLP project.

The SwiftKey's NLP project is mainly aimed to implement NLP(Natural language processing) techniques to build an algorithm in R environment. The algorithm will be trained by a larged amounts of collected text/documents and eventually, it will be able to make predictions of words that users are most likely to type. This predictive model will be built up for mobile keyboards, so a tradeoff between model capbility and efficiency will be considered during project.

The word prediction project has been started and an initial assessment of the task completed. This report describes the goals of the overall project, the magnitude of the challenge, and gives some initial assessments of the results.

In this report, the basic research like preprocess of raw data, preliminary statistics/visualization analysis, plans for algorithm and applications will be introduced so that reader can have an overall concept regarding the project and where the project is heading to.

Following flowchart gives us the main steps will be included of SwiftKey's NLP project. 


### Preprocess
The preprocess for text mining mainly includes cleaning, tokenization and stemming. The objectives of these processes are to clean the collections of text documents provided and transfer the documents into a form of text segmentation which can be used for further analysis easily. To be more specific, the following issues in text documents will be solved during preprocess:

1. Capital/Lower case
2. Numbers
3. Punctuations
4. Whitespace
5. Profanity words
6. Special notation/Noise like mistypes, UTF-16 encoded characters, foreign words, etc.

To overcome all issues above, function tokenization() has been constructed and following is an output of applying this function on our documents.
 
```{r echo=FALSE}
setwd('/Users/ivan/Work_directory/SwiftKey')
en_US <- file.path('.','final','en_US_s')
en_US.document <- Corpus(DirSource(en_US, encoding="UTF-8"), 
                         readerControl = list(reader = readPlain,language = "en_US",load = TRUE))
docs <- en_US.document
trans <- c(F,T,T,T,F,F,T,T)
ChartoSpace <- c('/','\\|')
stopWords <- 'english'
ownStopWords <- c()
swearwords <- read.table('SwiftKey-Natural-language/profanity filter/en', sep='\n')
names(swearwords)<-'swearwords'
filter <- rep('***', length(swearwords))
profanity <- data.frame(swearwords, target = filter)
profanity <- rbind(profanity, data.frame(swearwords = c("[^[:alpha:][:space:]']","â ","ã","ð"), target = c(" ","'","'","'")))
tokenized_docs <- tokenization(docs, trans, ChartoSpace, stopWords, ownStopWords, profanity)
```

After tokenizations, stemming is applied to documents to remove common words endings for English words, such as "es", "ed" and "s". 

```{r echo=FALSE}
stem_docs <- tm_map(tokenized_docs, stemDocument, 'english') # SnowballStemmer
```

So far, we have done the basic cleaning/transformation steps for raw documents. Next we will do some preliminary statistics analysis and data visualisation to collected data.

### Preliminary Statistics/Visualization
In this part, through doing some basic statistics analysis and data visualization on our data sets, we can get a brief understanding of our data. 
First, we explore the total lines and number of words in each document.

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=4}
setwd('/Users/ivan/Work_directory/SwiftKey')
load('./SwiftKey-Natural-language/Other/melt_df.RData')
melt_df
# barplot of total word counts
ggplot(melt_df,aes(x=names,y=value)) + 
    geom_bar(stat='identity',color='grey60',fill='cornsilk') + facet_wrap(~variable, scales = "free") + 
    xlab('Source')+ylab('Word Count/Lines') + theme(legend.position='none') + 
    ggtitle('Total Word Count / Lines by Text Source') + theme_bw() +
    geom_text(aes(label=format(value,big.mark=","),size=3),labels=comma,vjust=-0.2) 
```

Second, we convert our text corpus into Document Term Matrix based on different ngrams, so that we can easily figure out the frequency and correlation between different words. 

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=4}
setwd('/Users/ivan/Work_directory/SwiftKey')
rm(list=ls(all=TRUE));gc(reset=TRUE);par(mfrow=c(1,1))
source('SwiftKey-Natural-language/Task_1.5_Tokenization_func.R')
load('./data/all_corpus.RData')
```

sample
corpus - data.frame - ngram

ngram 1-4
words DTM (keep)
correlation diagram
word cloud

df <- data.frame(text=unlist(sapply(stem_docs, '[',"content")),stringsAsFactors=F)
df<-df[regexpr(pattern = '^([a-zA-Z])(?!(\\1{1,}))[a-zA-Z]*([a-zA-Z]+-([a-zA-Z]){2,})?(\'(s)?)?$', df, perl=T )>0]

### Prediction Algorithm
sparsity

### Application
After training the predictive algorithm, we also need to develop an online application with user-friendly interface. In this project, the application will be settled on a Shiny server. However, considering that the algorithm is built for mobile app, we have to also take the size and speed of model into account. So we only implement 1-4 grams algorithms for our online application.

The main functionalities should be included in Shiny app:

1. Detecting the nearest 1 to 3 words of users' typing and taking them as the inputs of model. 
2. Return the predictions of model to the user interface.

Example: 
user input = “allows you” 
predicted word = “to”

